The Concurrent Calculi Formalisation Benchmark

### 2024-02-19 short-paper/short-paper.tex
Edit intro part of design section to integrate design goals.

Our benchmark problems (detailed in \cref{app:challenges}) consider
\emph{in isolation} three key aspects that frequently pose challenges
when mechanising concurrency theory: \emph{linearity}, \emph{scope
  extrusion}, and \emph{coinductive reasoning}.  Mechanisations must
often address several of these aspects at the same time; however, we
see the combination of techniques as a next
step%beyond the scope of this challenge
, as discussed further in \Cref{sec:going-beyond}.  Our benchmark
challenges are based on process calculi, as these provide a simple but
realistic showcase of common issues encountered during mechanisation
efforts.
% While the three highlighted aspects are not the sole sources of difficulties, they emerge in most mechanisations of concurrency theory.

@@

\section{Design of the Benchmark}\label{sec:design-discussion}
In this section, we outline the factors considered in designing the benchmark challenges.
We begin with some general remarks, then describe the individual design considerations for each challenge problem, and the evaluation criteria for solutions.

Similarly to the authors of the POPLMark challenges, we seek to
answer several questions:
\alb{Should we stress we are interested in the \textbf{meta-theory} of process calculi, not}
\begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
\item\label{item:rq1} What is the current state of the art in mechanising process calculi?
\item\label{item:rq2} Which techniques and best practices can be recommended when starting formalisation projects involving process calculi?
\item\label{item:rq3} Which improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
\end{enumerate}

To provide a framework in which to answer these questions, our benchmark is designed to satisfy three main goals:
\begin{enumerate}[label=\textbf{(G\arabic*)},leftmargin=10mm]
\item\label{item:goal-comperison-accessibility} To enable the comparison of
  proof mechanisation approaches, making the challenges accessible to
  mechanisation experts who may be unfamiliar with concurrency theory;

@@

\todo{Refer to the future benchmark section about other things one could do, but they still all need the basics}
Still, these (and other) applications need the basic
techniques that our challenge problems emphasise:
Aligned with design goal~\ref{item:goal-reusability}, our problems are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory. These problems centre around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in future endeavours.
To achieve design goal \ref{item:goal-comperison-accessibility}, we have
formulated our challenge problems to explore the three aspects
(linearity, scope extrusion, coinduction) independently, so that they
may be solved individually and in any order; each problem is
reasonably small and easily understandable with basic knowledge of
textbook concurrency theory, process calculi, and type theory.  The
process calculus used in each challenge focuses on the features that
we want to emphasise, and omits all constructs (such as choices) that
would complicate the mechanisation without bringing additional
tangible benefits and insights.  The minimality and uniformity of the
calculi also allows us to target design goal
\ref{item:goal-tutorials}.

>>>>

Our benchmark problems (detailed in \cref{app:challenges}) consider
\emph{in isolation} three key aspects that frequently pose challenges
when mechanising concurrency theory: \emph{linearity}, \emph{scope
  extrusion}, and \emph{coinductive reasoning}.  Mechanisations must
often address several of these aspects at the same time; however, we
see the combination of techniques as a next
step%beyond the scope of this challenge
, as discussed further in \cref{sec:going-beyond}.  Our benchmark
challenges are based on process calculi, as these provide a simple but
realistic showcase of common issues encountered during mechanisation
efforts.
% While the three highlighted aspects are not the sole sources of difficulties, they emerge in most mechanisations of concurrency theory.

@@

\section{Design of the Benchmark}\label{sec:design-discussion}
In this section, we outline our considerations in designing the benchmark challenges.
We will begin with our design goals, then describe design considerations for each challenge problem, and finally suggest some evaluation criteria for solutions.

Similarly to the authors of the POPLMark challenges, we seek to
answer several questions about mechanising the meta-theory of process calculi:
\begin{enumerate}[label=\textbf{(Q\arabic*)},leftmargin=10mm]
\item\label{item:rq1} What is the current state of the art in tools for mechanising the meta-theory of process calculi?
\item\label{item:rq2} Which techniques and best practices can be recommended when starting new mechanisation projects involving process calculi?
\item\label{item:rq3} Which improvements are needed to make mechanisation tools more user-friendly with regards to the issues faced when mechanising process calculi?
\end{enumerate}

To provide a framework in which to answer these questions, our benchmark is designed to satisfy three main design goals:
\begin{enumerate}[label=\textbf{(G\arabic*)},leftmargin=10mm]
\item\label{item:goal-comperison-accessibility} To enable the comparison of
  proof mechanisation approaches by making the challenges accessible to
  mechanisation experts who may be unfamiliar with concurrency theory;

@@

To achieve design goal \ref{item:goal-comperison-accessibility}, we have
formulated our challenge problems to explore the three aspects
(linearity, scope extrusion, coinduction) independently, so that they
may be solved individually and in any order; each problem is
reasonably small and easily understandable with basic knowledge of
textbook concurrency theory, process calculi, and type theory.  The
process calculus used in each challenge focuses on the features that
we want to emphasise, and omits all constructs (such as choices) that
would complicate the mechanisation without bringing additional
tangible insights.  The minimality and uniformity of the
calculi also allows us to target design goal
\ref{item:goal-tutorials}.
Aligned with design goal~\ref{item:goal-reusability}, our problems are rooted in the fundamental meta-theory of process calculi, without necessitating the creation of novel theory.
Our challenges centre around well-established theorems and results, showcasing interesting proof techniques that can be leveraged in many applications (as we will further discuss in \cref{sec:going-beyond}).
